Tackling discrepancies in freshwater invertebrate trait databases: Harmonising across continents and aggregating taxonomic resolution

### 2021-07-14 Paper/Submission_new/Latex_Supplementary_File.tex
14.07: Revisions change in harm. first part

We selected the ecologically relevant traits identified by \citet{szocs_effects_2014} that increased (shredder, life cycle duration $ > 1 $ year, gills, bi/multivoltine, ovoviviparity) or decreased (univoltine, eggs in clutches/cemented or fixed, life cycle duration $ <= 1$ year) with increasing salinity and used these in the fourth corner analysis. The selected traits have also reacted to salinity in other studies (but see the discussion in \citet{szocs_effects_2014}). 
For some sites, environmental and abundance data have been sampled multiple times per year. \citet{szocs_effects_2014} used the full data and did not consider repeated sampling. We applied two analysis: one where we did not consider multiple samplings. And second where we selected, for the sites with multiple samplings, only those sampling events with the highest macroinvertebrate abundances per year and accordingly only used the conductivity data for these sampling events (this was not done for the RDA approach).Fourth corner analysis was carried out in R with the \textit{fourthcorner()} function from the ade4 package.
With the fourth corner analysis fewer trait-conductivity relationships than with the RDA approach were detected. With the original dataset significant relationships to conductivity were found for the traits shredder and ovoviviparity. In three out of the six aggregated and harmonised datasets (\textit{direct\_agg\textsubscript{mean}}, \textit{weighted\_agg}, harmonised (not aggregated)) four to five significant relationships (for the traits shredder, univoltinism, bi/multivoltine, ovoviviparity, and aquatic oviposition) were detected. In the remaining three datasets (\textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{direct\_agg\textsubscript{median}}) two significant relationships for the traits univoltinism and bi/multivoltinism were detected. Results were only slightly different when not considering multiple samplings (original dataset: one significant trait-conductivity relationship, others three to four). For the significant trait-conductivity relationships, the classification into characteristic traits for downstream and upstream sites was consistent with the results of the RDA analysis.

>>>>

We selected the ecologically relevant traits identified by \citet{szocs_effects_2014} that increased (shredder, life cycle duration $ > 1 $ year, gills, bi/multivoltine, ovoviviparity) or decreased (univoltine, eggs in clutches/cemented or fixed, life cycle duration $ <= 1$ year) with increasing salinity and used these in the fourth corner analysis. The selected traits have also reacted to salinity in other studies (but see the discussion in \citet{szocs_effects_2014}).

For some sites, environmental and abundance data have been sampled multiple times per year. \citet{szocs_effects_2014} used the full data and did not consider repeated sampling. We applied two analysis: one where we did not consider multiple samplings. And second where we selected, for the sites with multiple samplings, only those sampling events with the highest macroinvertebrate abundances per year and accordingly only used the conductivity data for these sampling events (this was not done for the RDA approach).Fourth corner analysis was carried out in R with the \textit{fourthcorner()} function from the ade4 package.

Fourth corner analysis detected fewer trait-conductivity relationships than the RDA approach. With the original dataset significant relationships to conductivity were found for the traits shredder and ovoviviparity. In three out of the six aggregated and harmonised datasets (\textit{direct\_agg\textsubscript{mean}}, \textit{weighted\_agg}, \textit{direct\_agg\textsubscript{median}}) four significant relationships (for the traits univoltinism, bi/multivoltine, ovoviviparity, and aquatic oviposition) were detected. In the remaining three datasets (\textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and harmonised (not aggregated)) two significant relationships for the traits univoltinism and bi/multivoltinism were detected (Table \ref{stab:fc_selected_reduced}). Results were only slightly different when not considering multiple samplings (original dataset: two significant trait-conductivity relationship (bi-multivoltine, ovoviviparity), other datasets two to four, Table \ref{stab:fc_selected_full}). For all significant trait-conductivity relationships, the classification into characteristic traits for downstream and upstream sites was consistent with the results of the RDA analysis.

\begin{longtable}[H]{m{2.6cm}|m{7.3cm}|m{1cm}|m{1.4cm}|m{1.5cm}|m{1.5cm}}
    \caption{Results of the fourth corner analysis, after accounting for multiple samplings per year, with the original dataset, harmonised, and aggregated trait datasets.}
    \label{stab:fc_selected_reduced}
    Dataset & Test & Obs & Std.Obs & Pvalue & Pvalue.adj \\ 
       stepw\_median & cond / feed\_shredder & 0.45 & 1.86  & 0.04 & 0.16 \\ 
       stepw\_median & cond / resp\_gil & 0.47 & 1.87  & 0.05 & 0.16 \\ 
       stepw\_median & cond / volt\_uni & -0.57 & -2.29  & 0.00 & 0.03 \\ 
       stepw\_median & cond / volt\_bi\_multi & 0.58 & 2.35  & 0.00 & 0.02 \\ 
       stepw\_median & cond / ovip\_aqu & -0.53 & -2.14  & 0.01 & 0.06 \\ 
       stepw\_median & cond / ovip\_ovo & 0.54 & 2.15  & 0.01 & 0.06 \\ 
       stepw\_median & cond / Life cycle duration $<=$ 1 year & -0.32 & -1.27  & 0.24 & 0.49 \\ 
       stepw\_median & cond / Life cycle duration $>$ 1 year & 0.32 & 1.27  & 0.24 & 0.49 \\ 
       stepw\_mean & cond / feed\_shredder & 0.46 & 1.96  & 0.03 & 0.11 \\ 
       stepw\_mean & cond / resp\_gil & 0.47 & 1.75  & 0.06 & 0.18 \\ 
       stepw\_mean & cond / volt\_uni & -0.58 & -2.45  & 0.01 & 0.04 \\ 
       stepw\_mean & cond / volt\_bi\_multi & 0.59 & 2.50  & 0.00 & 0.03 \\ 
       stepw\_mean & cond / ovip\_aqu & -0.52 & -2.23  & 0.01 & 0.06 \\ 
       stepw\_mean & cond / ovip\_ovo & 0.52 & 2.23  & 0.01 & 0.06 \\ 
       stepw\_mean & cond / Life cycle duration $<=$ 1 year & -0.33 & -1.30  & 0.23 & 0.47 \\ 
       stepw\_mean & cond / Life cycle duration $>$ 1 year & 0.33 & 1.30  & 0.23 & 0.47 \\ 
       direct\_median & cond / feed\_shredder & 0.45 & 1.89  & 0.04 & 0.15 \\ 
       direct\_median & cond / resp\_gil & 0.47 & 1.88  & 0.04 & 0.15 \\ 
       direct\_median & cond / volt\_uni & -0.58 & -2.24  & 0.00 & 0.02 \\ 
       direct\_median & cond / volt\_bi\_multi & 0.58 & 2.33  & 0.00 & 0.02 \\ 
       direct\_median & cond / ovip\_aqu & -0.53 & -2.14  & 0.01 & 0.05 \\ 
       direct\_median & cond / ovip\_ovo & 0.54 & 2.15  & 0.01 & 0.05 \\ 
       direct\_median & cond / Life cycle duration $<=$ 1 year & -0.34 & -1.35  & 0.21 & 0.43 \\ 
       direct\_median & cond / Life cycle duration $>$ 1 year & 0.34 & 1.35  & 0.21 & 0.43 \\ 
       direct\_mean & cond / feed\_shredder & 0.46 & 1.83  & 0.05 & 0.19 \\ 
       direct\_mean & cond / resp\_gil & 0.47 & 1.82  & 0.05 & 0.19 \\ 
       direct\_mean & cond / volt\_uni & -0.58 & -2.49  & 0.00 & 0.02 \\ 
       direct\_mean & cond / volt\_bi\_multi & 0.59 & 2.54  & 0.00 & 0.02 \\ 
       direct\_mean & cond / ovip\_aqu & -0.52 & -2.26  & 0.01 & 0.05 \\ 
       direct\_mean & cond / ovip\_ovo & 0.52 & 2.43  & 0.01 & 0.05 \\ 
       direct\_mean & cond / Life cycle duration $<=$ 1 year & -0.34 & -1.37  & 0.18 & 0.37 \\ 
       direct\_mean & cond / Life cycle duration $>$ 1 year & 0.34 & 1.37  & 0.18 & 0.37 \\ 
       weighted & cond / feed\_shredder & 0.46 & 1.97  & 0.03 & 0.11 \\ 
       weighted & cond / resp\_gil & 0.47 & 1.92  & 0.04 & 0.11 \\ 
       weighted & cond / volt\_uni & -0.58 & -2.51  & 0.00 & 0.02 \\ 
       weighted & cond / volt\_bi\_multi & 0.59 & 2.56  & 0.00 & 0.02 \\ 
       weighted & cond / ovip\_aqu & -0.52 & -2.13  & 0.01 & 0.04 \\ 
       weighted & cond / ovip\_ovo & 0.52 & 2.35  & 0.01 & 0.04 \\ 
       weighted & cond / Life cycle duration $<=$ 1 year & -0.34 & -1.38  & 0.20 & 0.41 \\ 
       weighted & cond / Life cycle duration $>$ 1 year & 0.34 & 1.38  & 0.20 & 0.41 \\ 
       not\_aggregated & cond / feed\_shredder & 0.51 & 2.08  & 0.01 & 0.08 \\ 
       not\_aggregated & cond / resp\_gil & 0.46 & 1.72  & 0.08 & 0.40 \\ 
       not\_aggregated & cond / volt\_uni & -0.64 & -2.76  & 0.00 & 0.02 \\ 
       not\_aggregated & cond / volt\_bi\_multi & 0.65 & 2.81  & 0.00 & 0.02 \\ 
       not\_aggregated & cond / ovip\_aqu & -0.26 & -1.22  & 0.27 & 0.80 \\ 
       not\_aggregated & cond / ovip\_ovo & 0.27 & 1.39  & 0.20 & 0.80 \\ 
       not\_aggregated & cond / Life cycle duration $<=$ 1 year & -0.33 & -1.26   & 0.24 & 0.80 \\ 
       not\_aggregated & cond / Life cycle duration $>$ 1 year & 0.33 & 1.26  & 0.24 & 0.80 \\ 
       original & cond / Life cycle duration $<=$ 1 year & -0.33 & -1.31  & 0.22 & 0.65 \\ 
       original & cond / Life cycle duration $>$ 1 year & 0.33 & 1.31  & 0.22 & 0.65 \\ 
       original & cond / Potential number of cycles per year $1$ & -0.48 & -1.91  & 0.04 & 0.18 \\ 
       original & cond / Potential number of cycles per year $> 1$  & 0.49 & 1.95  & 0.03 & 0.16 \\ 
       original & cond / Reproduction ovoviviparity & 0.58 & 2.43  & 0.01 & 0.04 \\ 
       original & cond / Reproduction clutches, cemented or fixed & -0.33 & -1.27  & 0.24 & 0.65 \\ 
       original & cond / Respiration gill & 0.47 & 1.87  & 0.04 & 0.18 \\ 
       original & cond / Feeding habits shredder & 0.54 & 2.51  & 0.01 & 0.04 \\ 

\begin{longtable}[H]{m{2.6cm}|m{7.3cm}|m{1cm}|m{1.4cm}|m{1.5cm}|m{1.5cm}}
    \caption{Results of the fourth corner analysis without accounting for multiple samplings per year, with the original dataset, harmonised, and aggregated trait datasets.}
    \label{stab:fc_selected_full}
    Dataset & Test & Obs & Std.Obs & Pvalue & Pvalue.adj \\ 
    stepw\_median & cond / feed\_shredder & 0.38 & 1.86 & 0.05 & 0.19 \\ 
      stepw\_median & cond / resp\_gil & 0.38 & 1.80 & 0.06 & 0.19 \\ 
      stepw\_median & cond / volt\_uni & -0.54 & -2.43 & 0.00 & 0.03 \\ 
      stepw\_median & cond / volt\_bi\_multi & 0.54 & 2.50 & 0.00 & 0.02 \\ 
      stepw\_median & cond / ovip\_aqu & -0.46 & -2.34 & 0.01 & 0.04 \\ 
      stepw\_median & cond / ovip\_ovo & 0.46 & 2.57 & 0.01 & 0.04 \\ 
      stepw\_median & cond / Life cycle duration $<=$ 1 year & -0.28 & -1.29 & 0.22 & 0.45 \\ 
      stepw\_median & cond / Life cycle duration $>$ 1 year & 0.28 & 1.29 & 0.22 & 0.45 \\ 
      stepw\_mean & cond / feed\_shredder & 0.39 & 1.93 & 0.03 & 0.12 \\ 
      stepw\_mean & cond / resp\_gil & 0.39 & 1.81 & 0.06 & 0.17 \\ 
      stepw\_mean & cond / volt\_uni & -0.54 & -3.65 & 0.00 & 0.01 \\ 
      stepw\_mean & cond / volt\_bi\_multi & 0.55 & 3.71 & 0.00 & 0.01 \\ 
      stepw\_mean & cond / ovip\_aqu & -0.44 & -2.26 & 0.01 & 0.05 \\ 
      stepw\_mean & cond / ovip\_ovo & 0.45 & 2.50 & 0.00 & 0.02 \\ 
      stepw\_mean & cond / Life cycle duration $<=$ 1 year & -0.28 & -1.34 & 0.20 & 0.39 \\ 
      stepw\_mean & cond / Life cycle duration $>$ 1 year & 0.28 & 1.34 & 0.20 & 0.39 \\ 
      direct\_median & cond / feed\_shredder & 0.37 & 1.89 & 0.03 & 0.13 \\ 
      direct\_median & cond / resp\_gil & 0.38 & 1.82 & 0.05 & 0.16 \\ 
      direct\_median & cond / volt\_uni & -0.54 & -3.79 & 0.00 & 0.01 \\ 
      direct\_median & cond / volt\_bi\_multi & 0.55 & 3.85 & 0.00 & 0.01 \\ 
      direct\_median & cond / ovip\_aqu & -0.46 & -2.37 & 0.01 & 0.05 \\ 
      direct\_median & cond / ovip\_ovo & 0.46 & 2.56 & 0.01 & 0.05 \\ 
      direct\_median & cond / Life cycle duration $<=$ 1 year & -0.30 & -1.41 & 0.18 & 0.35 \\ 
      direct\_median & cond / Life cycle duration $>$ 1 year & 0.30 & 1.41 & 0.18 & 0.35 \\ 
      direct\_mean & cond / feed\_shredder & 0.39 & 1.94 & 0.03 & 0.11 \\ 
      direct\_mean & cond / resp\_gil & 0.38 & 1.79 & 0.06 & 0.17 \\ 
      direct\_mean & cond / volt\_uni & -0.55 & -3.54 & 0.00 & 0.01 \\ 
      direct\_mean & cond / volt\_bi\_multi & 0.55 & 3.60 & 0.00 & 0.01 \\ 
      direct\_mean & cond / ovip\_aqu & -0.44 & -2.25 & 0.01 & 0.03 \\ 
      direct\_mean & cond / ovip\_ovo & 0.45 & 2.52 & 0.00 & 0.02 \\ 
      direct\_mean & cond / Life cycle duration $<=$ 1 year & -0.30 & -1.49 & 0.15 & 0.30 \\ 
      direct\_mean & cond / Life cycle duration $>$ 1 year & 0.30 & 1.49 & 0.15 & 0.30 \\ 
      weighted & cond / feed\_shredder & 0.39 & 1.84 & 0.05 & 0.19 \\ 
      weighted & cond / resp\_gil & 0.38 & 1.81 & 0.06 & 0.19 \\ 
      weighted & cond / volt\_uni & -0.55 & -3.43 & 0.00 & 0.01 \\ 
      weighted & cond / volt\_bi\_multi & 0.55 & 3.49 & 0.00 & 0.01 \\ 
      weighted & cond / ovip\_aqu & -0.44 & -2.24 & 0.01 & 0.05 \\ 
      weighted & cond / ovip\_ovo & 0.45 & 2.44 & 0.01 & 0.05 \\ 
      weighted & cond / Life cycle duration $<=$ 1 year & -0.30 & -1.42 & 0.17 & 0.34 \\ 
      weighted & cond / Life cycle duration $>$ 1 year & 0.30 & 1.42 & 0.17 & 0.34 \\ 
      not\_aggregated & cond / feed\_shredder & 0.43 & 2.07 & 0.01 & 0.09 \\ 
      not\_aggregated & cond / resp\_gil & 0.38 & 1.77 & 0.06 & 0.30 \\ 
      not\_aggregated & cond / volt\_uni & -0.57 & -3.90 & 0.00 & 0.01 \\ 
      not\_aggregated & cond / volt\_bi\_multi & 0.58 & 3.96 & 0.00 & 0.01 \\ 
      not\_aggregated & cond / ovip\_aqu & -0.25 & -1.28 & 0.22 & 0.62 \\ 
      not\_aggregated & cond / ovip\_ovo & 0.25 & 1.50 & 0.15 & 0.62 \\ 
      not\_aggregated & cond / Life cycle duration $<=$ 1 year & -0.29 & -1.37 & 0.18 & 0.62 \\ 
      not\_aggregated & cond / Life cycle duration $>$ 1 year & 0.29 & 1.37 & 0.18 & 0.62 \\ 
      original & cond / Life cycle duration $<=$ 1 year & -0.29 & -1.42 & 0.17 & 0.50 \\ 
      original & cond / Life cycle duration $>$ 1 year & 0.29 & 1.42 & 0.17 & 0.50 \\ 
      original & cond / Potential number of cycles per year $1$ & -0.47 & -2.11 & 0.02 & 0.10 \\ 
      original & cond / Potential number of cycles per year $> 1$ 1 & 0.47 & 2.19 & 0.01 & 0.05 \\ 
      original & cond / Reproduction ovoviviparity & 0.51 & 3.06 & 0.00 & 0.01 \\ 
      original & cond / Reproduction clutches, cemented or fixed & -0.34 & -1.49 & 0.17 & 0.50 \\ 
      original & cond / Respiration gill & 0.38 & 1.80 & 0.06 & 0.25 \\ 
      original & cond / Feeding habits shredder & 0.45 & 2.19 & 0.01 & 0.06 \\ 

### 2021-07-14 Paper/Submission_new/Main_Document-Latex.tex
14.07: Revisions change in harm. first part

The calculation of FD metrics for the original, harmonised (not aggregated), and aggregated datasets suggested that harmonising trait data preserves FRic and FDiv values, while aggregated trait data only preserved FRic values. The FRic and FDiv values calculated from harmonised trait data correlated highly to FRic and FDiv values from the original dataset ($R^2 = 0.96$ and $0.86$, respectively). FRic and FDiv calculated for aggregated trait data showed moderate to good correlations to the FRic and FDiv values calculated with the original dataset (FRic: $0.78 <= R^2 <= 0.85$ , FDiv: $0.47 <= R^2 <= 0.55$, Figure \ref{fig:FRic} and \ref{fig:FDiv}). By contrast, FEve values calculated with harmonised and aggregated trait data were unrelated to the FEve values calculated with the original dataset (all $R^2 < 0.1$), except for \textit{stepwise\_agg \textsubscript{median}}, which showed a moderate correlation to FEve values calculated from the original dataset ($R^2  = 0.48$, Figure \ref{fig:FEve}).

@@

FD analysis is an important trait-based application to quantify the variability of traits that influence ecosystem functions. We found through correlation that harmonised trait data may be used to calculate FRic and FDiv values, and that aggregated trait data can at least be used to calculate FRic values. This does not apply to FEve, as the use of harmonised and aggregated trait data yielded completely different FEve values than with the trait data from \citet{szocs_effects_2014}. FRic represents the trait space spanned by the community (also termed the convex hull volume) and is only dependent on the trait data used. Our results show that harmonisation preserves the trait space the best, i.e. the most extreme trait values of the original study, and that aggregation only slightly changes the trait space. It follows that FDiv, which describes how abundances are distributed within the volume of the trait space, is best preserved by harmonised trait data preserved and to a lesser extent with aggregated trait data. FEve describes how regular the trait space is filled by taxa. We found that harmonised and aggregated trait change the evenness of the abundance distribution in the trait space, which is likely because during the calculation of FEve only the nearest neighbours in the trait space are considered. This leads to the undesirable property that even small community changes result in large changes in the FEve index, a behaviour that has been shown by \citet{botta-dukatTestingAbilityFunctional2016} through simulation.

>>>>

The calculation of FD metrics for the original, harmonised (not aggregated), and aggregated datasets suggested that harmonising trait data preserves FRic and FDiv values, while aggregated trait data only preserved FRic values. The FRic and FDiv values calculated from harmonised trait data correlated highly to FRic and FDiv values from the original dataset ($R^2 = 0.96$ and $0.86$, respectively). FRic and FDiv calculated for aggregated trait data showed moderate to good correlations to the FRic and FDiv values calculated with the original dataset (FRic: $0.78 <= R^2 <= 0.84$ , FDiv: $0.42 <= R^2 <= 0.62$, Figure \ref{fig:FRic} and \ref{fig:FDiv}). By contrast, FEve values calculated with harmonised and aggregated trait data were unrelated to the FEve values calculated with the original dataset (all $R^2 < 0.1$), except for \textit{stepwise\_agg \textsubscript{median}}, which showed a moderate correlation to FEve values calculated from the original dataset ($R^2  = 0.48$, Figure \ref{fig:FEve}).

@@

FD analysis is an important trait-based application to quantify the variability of traits that influence ecosystem functions. We found through correlation that harmonised trait data may be used to calculate FRic and FDiv values, and that aggregated trait data can at least be used to calculate FRic values. This does not apply to FEve, as the use of harmonised and aggregated trait data yielded completely different FEve values than with the trait data from \citet{szocs_effects_2014}. FRic represents the trait space spanned by the community (also termed the convex hull volume) and is only dependent on the trait data used. Our results show that harmonisation preserves the trait space the best, i.e. the most extreme trait values of the original study, and that aggregation only slightly changes the trait space. It follows that FDiv, which describes how abundances are distributed within the volume of the trait space, is best preserved by harmonised trait data and to a lesser extent with aggregated trait data. FEve describes how regular the trait space is filled by taxa. We found that harmonised and aggregated trait change the evenness of the abundance distribution in the trait space, which is likely because during the calculation of FEve only the nearest neighbours in the trait space are considered. This leads to the undesirable property that even small community changes result in large changes in the FEve index, a behaviour that has been shown by \citet{botta-dukatTestingAbilityFunctional2016} through simulation.
